The Hydrogen Epoch of Reionization Array Dish I: Beam Pattern Measurements and Science Implications

### 2015-12-26 DishBeamMeasurements/paper/DishBeamMeasurementsPaper.tex
keep pushing

\section{Dish Design and Modeling}

@@

ALSO MENTION DEPENDENCE ON BASELINE DIRECTION
DISCUSS DELAY CLEANING TO THE 2min THERMAL NOISE LEVEL, THEN DELAY TRANSFORM AGAIN WITH REDUCED RESOLUTION, MEANING A LARGER BH WINDOW, AND THUS MORE LEAKAGE, THESE EFFECTS ARE DISCUSSED IN NITHYA'S PAPER, here we are focused on intrinsic delay distribution of foregrounds

We study in this section the effects of the beam power pattern on the apparent frequency dependence of the foregrounds, and in particular the uncertainties in these predictions due to beam modeling uncertainties. Thyagarajan et al., (submitted)

We simulate the power spectra of foregrounds with different primary beam models under various observing conditions. Broadly, beam models with more response near the horizon see smooth spectrum sources with more frequency dependence (higher delay), appearing closer to the nominally clean EOR window, making it more at risk of leaking power inside. We use the per-baseline approach of XX and XX, simulating visibilities measured by specific baselines as a function of frequency, then computing the Fourier transform over frequency (the delay transform) and normalizing the result into a cosmological power spectrum \citep{perbaselinetechnique}.

We simulate visibilities for three beam models (HFSS, CST, and an Airy pattern for a 14\,m aperture) at various LSTs, modeling the sky as the sum of the Global Sky Model \citep{gsm} and a point source catalog (XX). We use a frequency spacing of 1\,MHz, sufficient to characterize delays within and just outside of the horizon limits on both baseline lengths we are concerned with, 14\,m and 42\,m. We use a total bandwidth of 80\,MHz (40\,MHz after applying the Blackman-Harris window) centered on 150\,MHz. This bandwidth is larger than the 10\,MHz thought to be safe from signal evolution over redshift, but is the bandwidth used in the power spectrum analysis of XX,XX to perform a high delay resolution empirical foreground subtraction (delay-space CLEANing) as described by XXX. We normalize the delay spectra into power spectra following   \citet{nithya15}.

Nithya et al (submitted) detail the implications of these foreground delay spectra for suppression strategies. We are concerned here with the role of the beam in these predictions, in particular what our beam model and its uncertainties imply about foreground suppression strategies. We begin by simulating delay spectra, plotted in Figure \ref{fig:delayspec} (top panel) at various LSTs as outlined above using the nominal HFSS beam. As all these LSTs are high galactic latitudes far from the galactic center, the total visibility power (the level of the zero delay mode) varies by only an order of magnitude over these LSTs both on the 14\,m (left panel) and 42\,m (right panel). However the horizon limit at negative delay (corresponding to the western horizon) shows a peak that varies by three orders of magnitude on the 14\,m baseline and two orders of magnitude on the 42\,m baseline, showing the stark difference in pitchfork brightening when the galaxy is just above versus just below the horizon.

In this figure we perform the approximate conversion from delay to k//, which we plot as a second x-axis at the top of the plot. For these short baselines, k// may be converted to k by adding kperp=0.005 or 0.02 in quadrature for these baselines at this redshift. These numbers are small compared to the range of k// plotted, and thus we interpret the k// axis as simply the k axis, and plot a 1D model power spectrum computed using 21cmFast as a dotted line for comparison.

\begin{figure*}[h]
\centering
\includegraphics[width=6in]{delay_spectra_dave_rich_airy_beams_manyLSTs.pdf}
\includegraphics[width=6in]{delay_spectra_dave_rich_airy_beams_LST2deg.pdf}
\includegraphics[width=6in]{delay_spectra_dave_rich_airy_beams_LST62deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspec}
\end{figure*}

We study the differences between our three beam models on two of these LSTs, one with maximal horizon brightening ($2^\circ$), and one with minimal horizon brightening ($62^\circ$). Figure \ref{fig:gsmplots} shows the sine-projected Global Sky Model in horizontal coordinates with units of Kelvin for these two LSTs. Dashed lines mark zenith angles 20$^\circ$, 40$^\circ$, 60$^\circ$, 80$^\circ$. These plots confirm that the large negative delay peak at both baseline's horizon limits at the 2 degree LST originates from the center of the galaxy being just above the horizon. In contrast several hours later, the galactic center is fully below the western horizon, leaving only a slight brightening near the eastern horizon due to the much weaker galactic anticenter.

\begin{figure*}[h]
\centering
\includegraphics[width=6in]{gsm_kelvin_LST_2deg_and_62deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:gsmplots}
\end{figure*}

We turn now to the differences in foreground delay spectra in these two observing conditions with different beam models. Figure \ref{fig:delayspec} (middle panel) shows the delay spectra for all three beam models at $2^\circ$ LST, which has the worst horizon brightening. We use both numerical beam models with 5\,m feed heights, as well as the Airy pattern for a 14\,m aperture. The delay spectra for these beams agree out to delays of 10--20\,ns on the 14\,m baseline and 30--40\,ns on the 62m baseline. These limits suggest that the beams track each other fairly well out to 15--20$^\circ$ from zenith, beyond which they cease to be correlated. This is essentially what is observed in Figure \ref{fig:measuredbeamslices} (middle panel). At larger delays, especially near the negative delay horizon limit, all three model delay spectra diverge due to the significant edge brightening which effectively discriminates between these models. The CST, HFSS, and Airy beams reach roughly -35\,dB, -38\,dB, and -50\,dB at the horizon (Figure \ref{fig:modelbeams}), consistent with the fact that the CST beam has the largest horizon brightening, followed by the HFSS beam, and then the Airy beam. This is seen in the delay spectra from both baseline lengths, though the edge brightening is much more significant in the longer baseline where it is less diluted by the zero delay emission.

In contrast, all three model agree much more closely when there is little or no edge brightening as in Figure \ref{fig:measuredbeamslices} (middle panel) where we plot the delay spectra for all three beams for 62$^\circ$ LST. There is still a modest edge brightening observed here, a flattening off near the horizon on the short baseline and a small peak on the long baseline, both due to the increasing solid angle per zenith angle at larger zenith angles. However as the emission at this LST at large delays is roughly the same temperature as emission from everywhere else on the sky, washing out the difference between the three beam models.

\section{Discussion}

%- power spectrum analyses by first generation instruments are ongoing, but face many challenges
%- that’s why HERA was designed to be optimized for 21cm

>>>>

\section{Dish Design and Modeling}
NEEDS TLC

@@

We turn in this section to the effects of the beam power pattern on the apparent frequency dependence of the foregrounds. Thyagarajan et al., (submitted) discuss the apparent frequency dependence of foregrounds in more detail as well as methods to mitigate to mitigate it such as delay space CLEANing. We focus on this section on the uncertainties in these foreground power spectrum simulations due to beam modeling uncertainties, but we must first discuss these foreground simulations themselves and their dependence on observing conditions.

We simulate foreground power spectra using different primary beam models at various local sidereal times (LSTs). Given that our measured dish power patterns agree well with both numerical models (HFSS and CST) in the main lobe but deviate in the sidelobes, and that these models make somewhat different assumptions about the dish surface, we take them as a representative pair of possible dish models. We use a feed height of 5\,m, a compromise between larger collecting area and smaller risk of coupling to adjacent dishes. We also include the Airy pattern for comparison as \citet{nithya15}. Beam models with weaker response near the horizon (such as the Airy pattern) downweight sources in this direction of high apparent frequency dependence. This reduces the magnitude of emission near the edge of the EOR window, reducing the risk it leaks inside. We use the per-baseline approach of \citet{parsons12a,parsons12b}, simulating visibilities measured by specific baselines as a function of frequency, then computing the Fourier transform over frequency (delay transform) and normalizing the result into a cosmological power spectrum following \citet{nithya15}.

In detail, we simulate visibilities for each beam model at various LSTs, modeling the sky as the sum of the Global Sky Model \citep{gsm} and the Culgoora \citep{Slee1995} and MWA Commissioning Survey \citep{MWACS} point source catalogs. We use a frequency spacing of 1\,MHz, sufficient to characterize delays within and just outside of the horizon limits on both baseline lengths we are concerned with, 14\,m and 42\,m. We use a total bandwidth of 100\,MHz (50\,MHz after applying the Blackman-Harris window) centered on 150\,MHz. This bandwidth is larger than the 10\,MHz thought to be safe from signal evolution over redshift, but is the bandwidth used in the wide band delay space foreground CLEAN of \citet{paper32,paper64}.

Figure \ref{fig:delayspec} (top panel) shows simulated foreground delay spectra at various LSTs using the nominal HFSS beam. As all these LSTs are high galactic latitudes far from the galactic center, the total visibility power (the level of the zero delay mode) varies only by a factor of a few over these LSTs on both baseline lengths (14\,m (left panel), 42\,m (right panel)). However the negative delay horizon limit (corresponding to the western horizon) has a peak that varies by over three orders of magnitude on the 14\,m baseline and by two orders of magnitude on the 42\,m baseline, demonstrating the stark difference in horizon brightening when the galaxy is just above versus just below the horizon.

In this figure we perform the approximate conversion from delay $\tau$ to $k_\parallel$, which we plot as a second $x$-axis at the top of the plot. For these short baselines, $k_\parallel$ may be converted to $k$ by adding $k_\perp=0.005$ for the 14\,m baseline or $k_\perp=0.02$ for the 42\,m baseline in quadrature. These numbers are small compared to the range of $k_\parallel$ plotted, and thus we interpret the $k_\parallel$ axis as simply the $k$ axis, and plot a 1D model power spectrum computed using 21cmFast \citep{21cmfast} as a dotted line for comparison.

\begin{figure*}[h]
\centering
\includegraphics[width=6in]{delay_spectra_dave_rich_airy_beams_manyLSTs.pdf}
\includegraphics[width=6in]{delay_spectra_dave_rich_airy_beams_LST2deg.pdf}
\includegraphics[width=6in]{delay_spectra_dave_rich_airy_beams_LST62deg.pdf}
\caption{Simulated foreground delay spectra using the HFSS beam at various LSTs (top panel). We plot the delay spectra at two LSTs spanning the range of possible horizon brightening for three different beam models. The maximum horizon brightening at the negative horizon occurs at 2$^\circ$ LST (middle panel), and the three beam models thus differ markedly in their predicted delay spectra near the horizon. In contrast, when the horizon brightening effect is smaller at 2$^\circ$ LST, the foreground delay spectra from all three beams agree much more closely.}
\label{fig:delayspec}
\end{figure*}

To characterize the effect of beam modeling uncertainties on this horizon brightening, we select two of these LSTs, one with maximal horizon brightening ($2^\circ$), and one with minimal horizon brightening ($62^\circ$). Figure \ref{fig:gsmplots} shows the sine-projected Global Sky Model, which dominates the horizon brightening effect, in horizontal coordinates with units of Kelvin for both LSTs. Dashed lines mark zenith angles 20$^\circ$, 40$^\circ$, 60$^\circ$, 80$^\circ$. These plots confirm that the large negative delay peak at the 2$^\circ$ LST is due to the center of the galaxy just above the horizon. In contrast several hours later, the galactic center is fully below the western horizon, leaving only a slight brightening near the eastern horizon due to the weaker galactic anticenter.

\begin{figure*}[h]
\centering
\includegraphics[width=6in]{gsm_kelvin_LST_2deg_and_62deg.pdf}
\caption{Global Sky Model \citep{gsm} in sine-projected horizontal coordinates at LST of 2$^\circ$ (left) and 62$^\circ$ right. The very bright emission from the center of the galaxy at the western horizon at 2$^\circ$ is seen in the delay spectra of EW baselines as a horizon brightening at negative delay.}
\label{fig:gsmplots}
\end{figure*}

How much do the predicted foreground power spectra differ between the three model dish power patterns? Figure \ref{fig:delayspec} (middle panel) shows the simulated delay spectra for all three beams at $2^\circ$ LST, when the horizon brightening is worst. All three delay spectra agree out to delays of 10--20\,ns on the 14\,m baseline and 30--40\,ns on the 40\,m baseline. These numbers suggest that the beams track each other fairly well out to 15--20$^\circ$ from zenith, beyond which they diverge. This is roughly what is observed in Figure \ref{fig:measuredbeamslices} (middle panel). At larger delays, especially near the negative delay horizon limit, all three model delay spectra diverge due to the significant edge brightening which effectively discriminates between these models. The CST, HFSS, and Airy beams reach roughly -35\,dB, -38\,dB, and -50\,dB at the horizon (Figure \ref{fig:modelbeams}), consistent with the fact that the CST beam has the largest horizon brightening, followed by the HFSS beam, and then by the Airy beam. This is seen in the delay spectra for both baseline lengths, though the edge brightening is much clearer on the longer baseline where it less diluted by zero delay emission.

In contrast, all three model agree much more closely when there is little or no edge brightening as in Figure \ref{fig:measuredbeamslices} (bottom panel) where we plot the delay spectra for all three beams for 62$^\circ$ LST. There is still a modest flattening off near the horizon on the 14\,m baseline and a slight peak on the 42\,m baseline due to the large solid angle near the horizon. However as the near horizon emission at this LST is roughly the same temperature as emission from everywhere else on the sky, the difference between the three beam models is greatly reduced.

\section{Discussion}
NEEDS TLC
%- power spectrum analyses by first generation instruments are ongoing, but face many challenges
%- that’s why HERA was designed to be optimized for 21cm
